---
title: "P2"
output: html_document
---
Please upload 4 things, as detailed in the Ellis & Leek paper:

-The raw data.
-A tidy data set (Wickham 2014).
-A code book describing each variable and its values in the tidy data set. [do this for all variables that will be part of your replication analyses]
-An explicit and exact recipe you used to go from 1 -> 2,3. [this must be an R script or R Markdown file]

I will upload: 
-this R Markdown file
-Study_1_Data.sav
- codebook...**************

-Study_1_Materials.doc

The raw data is contained in "Study_1_Data.sav". I brought this into the R environment using "Environment > Import Dataset > From SPSS". This file was obtained from "https://osf.io/9k3sw/", which is the data repository for the "Pain as Social Glue" paper, found at "https://journals.sagepub.com/doi/10.1177/0956797614545886". 

```{r}
library(tidyverse)
```

```{r}
glimpse(Study_1_Data)
```

2.2
3. There should be one table for each “kind” of data.

The "task" variables are ordinal, and they are in a good format (1:5 rankings, ascending). The variable names are sufficiently descriptive. 

The "feelings" variables are ordinal, and they are in a good format (1:5 rankings, ascending). The variable names are sufficiently descriptive. 

The "group" variables are ordinal, and they are in a good format (1:7 rankings, ascending). The variable names could be improved (e.g. group101 to group1 -- group question 1). 

```{r}
feel_col <- colnames(Study_1_Data)[startsWith(colnames(Study_1_Data), "feel")]
task_col <- colnames(Study_1_Data)[startsWith(colnames(Study_1_Data), "task")]
group_col <- colnames(Study_1_Data)[startsWith(colnames(Study_1_Data), "group")]
results_col <- c(colnames(Study_1_Data)[endsWith(colnames(Study_1_Data), "TOT")],colnames(Study_1_Data)[endsWith(colnames(Study_1_Data), "PANAS")])

demog <- Study_1_Data %>% select(subid,subnumber, subage, subgender, gendercoded, subbornaus, subbornloc, subausyears, subefl)
experiment <- Study_1_Data %>% select(subid, groupnumber, groupsize, condition, pain1secs, pain2secs)
task <- Study_1_Data %>% select(subid, task_col)
feel <- Study_1_Data %>% select(subid, feel_col)
group <- Study_1_Data %>% select(subid, group_col)
results <- Study_1_Data %>% select(subid, results_col)

glimpse(demog)
glimpse(experiment)
glimpse(task)
glimpse(feel)
glimpse(group)
glimpse(results)
```

codebook may be a Word document

include information about survey document from data website

A common format for this document is a Word file. There should be a section called
“Study design” that has a thorough description of the question being asked by the study
as well as how you collected the data. An additional section called “Code book” should be
provided to describe each variable and its units. This information is frequently conveyed
most simply in tabular form. In this case, the columns of the table would contain columns
including VariableName, Description, Units, CodingNotes, and OtherNotes. Further
columns that provide additional information to the statistician should be included. (Figure
1C)

Continuous variables are anything measured on a quantitative scale that could be any
fractional number. An example would be something like weight measured in kg. Ordinal
data are data that have a fixed, small (< 100) number of levels but are ordered. This could
be for example survey responses where the choices are: poor, fair, good. Categorical data
are data where there are multiple categories, but they aren’t ordered. One example would
be sex: male or female. Missing data are data that are unobserved and you don’t know
the mechanism. Missing values should be coded as NA. If, however, missingness is coded
in an alternative manner, this should be explicitly noted in the code book. Censored data
are data where you know the missingness mechanism on some level. Common examples
are a measurement being below a detection limit or a patient being lost to follow-up. They
should also be coded as NA when you don’t have the data. But you should also add a new
column to your tidy data called, “VariableNameCensored” which should have values of
TRUE if censored and FALSE if not. In the code book you should explain why those values
are missing. It is absolutely critical to report to the analyst if there is a reason you know
about that some of the data are missing. You should also not impute, make up, or throw
away missing observations.

Generally, try to avoid coding categorical or ordinal variables as numbers. When you
enter the value for sex in the tidy data, it should be “male” or “female”. The ordinal values
in the data set should be “poor”, “fair”, and “good” not 1, 2, 3. This coding is attractive
because it is self-documenting; any ambiguity or need for interpretation by the analyst is
removed. This will ultimately avoid potential mix-ups about which direction effects go and
will help identify coding errors.

## System Information
I developed and ran this code on Ubuntu Linux.
R Version: 3.4.4
RStudio Version: 1.1.456
Ubuntu Version: 18.04.1 LTS


whether you tried it more than once to confirm it gave the same results.